---
title: "ORFquant html report"
author: "Generated with ORFquant"
date: '`r format(Sys.time(), "%d %B, %Y")`'
params:
  input_files: ""
  input_sample_names: ""
  output_fig_path: ""
output:
  html_document: 
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ORFquant)
library(knitr) 
library(ggplot2) # plotting
library(reshape2) # data formatting for ggplot2
library(gridExtra) # multiple plots per figure
```

```{r data, echo=FALSE, warning=FALSE}
# get input from params
input_files <- params$input_files # RData file paths to one or more samples (one RData file per sample)
input_sample_names <- params$input_sample_names # user-defined sample names to be displayed in report (short(!) names)
names(input_files) <- input_sample_names
rdata_list<-list()
for(samplo in names(input_files)){
    rdata_list[[samplo]]<-get(load(input_files[[samplo]]))
}


```


# Data information

Sample names and data file paths visualized in this report:

```{r data_info, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
cat("\n\n")
for (i in 1:length(input_files)){
    cat("**", names(input_files[i]), "**: \n", sep="")
    cat(input_files[i], "\n\n\n")
}
cat("\n\n")
```


# 1 Transcript selection

As a first step in the *ORFquant* pipeline, Ribo-seq data on exonic bins and splice junctions is used to filter a subset of transcripts which can explain all the regions with signal, trying to minimize the number of transcripts with "empty" regions.\


In this section we can visualize the results from the filtering step.

## 1.1 Number of selected transcript per gene {.tabset}
```{r sel_1, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot barplot for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["Sel_txs_genes"]])
    cat("\n\n")
}
```

## 1.2 Number of selected transcripts per gene (against Ribo-seq TPM) {.tabset}

```{r sel_2, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot barplot for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["Sel_txs_genes_tpm"]])
    cat("\n\n")
}
```

## 1.3 % of selected transcripts per gene {.tabset}

```{r sel_3, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot barplot for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["Sel_txs_genes_pct"]])
    cat("\n\n")
}
```

## 1.4 Improvement on transcript-specific mapping {.tabset}

Results are shown using all annotated transcripts, transcripts annotated as protein-coding, and selected transcripts.
A covered features is a features with at least 1 P-site (or spliced read) mapping to it. 1 means uniquely mapping to a transcript, while 0 means mapping to none.

```{r sel_4, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.height=8,fig.width=10, dpi=150}
# plot barplot for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["Sel_txs_bins_juns"]])
    cat("\n\n")
}
```

# 2 Summary of detected ORFs

In this section summary statistics over ORF categories in different gene biotypes are shown.
**Note:** The definition and accuracy of ORF categories and biotypes depends also on the annotation used.\  

ORF categories:\  


**ORF_annotated**: ORF boundaries identical to annotation (start and stop codons).\  

**N_truncation**: downstream start codon but same stop codon as annotated.\  

**uORF**: start and stop codons upstream of annotated start codon.\  

**other**: overlapping uORF, nested ORFs, C- and N-extensions, dORF etc..\  

**not_annotated**: ORF in transcript with no CDS annotation.\  


Biotypes:\  

**Protein_coding**: transcript with annotated CDS.\  

**non-coding isoform**: non-coding transcript from protein-coding genes.\  

**pseudogene**: transcript from pseudogene (processed, unprocessed etc..).\  

**non-coding RNA**: non-coding transcript from non-coding gene (excluding pseudogenes).\  


Quantification metrics:\  

**ORFs_pM** (or ORFs per Million, akin to *TPM* for RNA-seq) indicates length-normalized abundance.\  

**ORF_pct_P_sites** (or % of gene translation) indicates the % of total translation for each gene. A value of 70 means that one ORF contains 70% of the total gene translation output (translation of other ORFs for that gene amounts to the remaining 30%).\


## 2.1 Number of detected ORFs {.tabset}

```{r norf_1, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_found"]])
    cat("\n\n")
}
```


## 2.2 % of gene translation {.tabset}

```{r norf_2, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_found_pct_tr"]])
    cat("\n\n")
}
```


## 2.3 ORF length {.tabset}

```{r norf_3, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_found_len"]])
    cat("\n\n")
}
```


## 2.4 ORFs per Million {.tabset}

```{r norf_4, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_found_ORFs_pM"]])
    cat("\n\n")
}
```


# 3 ORFs per gene

In this section general statistics about ORF-specific translation estimates to overall gene translational output are shown.\  
The *major ORF* is defined as the ORF harboring the highest number of P-sites positions (not normalized by ORF length).

## 3.1 Number of ORFs per gene {.tabset}

```{r orfg_1, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_genes"]])
    cat("\n\n")
}
```


## 3.2 Number of ORFs per gene (against Ribo-seq signal) {.tabset}

```{r orfg_2, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_genes_tpm"]])
    cat("\n\n")
}
```

## 3.3 Contribution from the major ORF per gene {.tabset}

```{r orfg_3, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_maxiso"]])
    cat("\n\n")
}
```

## 3.4 Contribution from the major ORF per gene (against Ribo-seq signal) {.tabset}

```{r orfg_4, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=10, dpi=150}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    print(rdata_list[[i]][["ORFs_maxiso_tpm"]])
    cat("\n\n")
}
```

## 3.5 Alternative splice site profiles {.tabset}

Optional, available if coverage files were provided: aggregate plot of Ribo-seq coverage (normalized 0-1 per each region) and ORF coverage over candidate alternative splice sites regions.\ No mixture indicates one ORF only, while other tracks indicate the presence of additional ORFs, divided by their summed translation values. Explanatory scheme at the bottom, with blue representing the major ORF and red the additional ORFs. Only regions with enough signal and separation from the major ORF are shown.

```{r orfg_5, echo=FALSE, warning=FALSE, message=FALSE, results="asis", fig.width=27,fig.height=16,dpi=300}
# plot distribution for each sample in separate tabs
for (i in c(1:length(rdata_list))){
    cat("### ", names(rdata_list)[i], " {.tabset} \n \n")
    aaa<-rdata_list[[i]][["Meta_splicing_coverage"]]
    if(length(aaa)>0){
        print(aaa)
    }
    cat("\n\n")
}
```

